Ensemble Stability of Biological Networks

### 2014-09-23 paper.tex
add error estimate for Monte Carlo simulation

Even though the determination of stability only requires examining
linearized equations of motion, it is worth noting that knowing the
stability of a fixed point can yield information about the non-linear
dynamics of a system.  For instance, in two dimensions,
Poincare-Bendixson theory implies that every limit cycle must encircle
at least one fixed point.  Furthermore, in the case where the cycle
contains a single fixed point (such as in the Volterra-Lotke model of
conflicting populations), the stability of the fixed point will be
determined by whether the limit cycle is attractive or repulsive.

@@

\todo{Say a few words about Monte Carlo integration and error bars.}

>>>>

Even though the determination of stability only requires examining
linearized equations of motion, it is worth noting that knowing the
stability of a fixed point can yield information about the non-linear
dynamics of a system.  For instance, in two dimensions,
Poincare-Bendixson theory implies that every limit cycle must encircle
at least one fixed point \cite{Davis1962}.  Furthermore, in the case where the cycle contains a single fixed point (such as in the Lotka-Volterra model of
conflicting populations), the stability of the fixed point will be
determined by whether the limit cycle is attractive or repulsive.

@@

If we sample $N = N_{\mathrm{stab}} + N_{\mathrm{unstab}}$ matrices where each has some probability $\theta$ of being stable then $N_{\mathrm{stab}}$ has a binomial distribution. We can compute a sample estimate for $\theta$, $\hat{\theta} = \frac{N_{\mathrm{stab}}}{N}$ \cite{Murphy2012}. The posterior distribution in this case is known to be a Beta distribution as a result of Beta-Binomial conjugacy
$$
\mathrm{Beta}(\theta | \mathcal{D}) = \mathrm{Beta}(\theta | N_{\mathrm{stab}} + a, N_{\mathrm{unstab}} + b)
$$
where $a$ and $b$ are the hyperparameters of the Beta prior and we consider the uninformative uniform prior corresponding to $a=b=1$. We consider the maximum a posteriori estimate
$$\hat{\theta}_{MAP} = \frac{a + N_{\mathrm{stab}} - 1}{a + b + N - 2}$$
which corresponds in this case to the maximum likelihood estimate
$$
\hat{\theta}_{MLE} = \frac{N_{\mathrm{stab}}}{N}.
$$
This estimate is characterized by the variance of the posterior Beta distribution
$$
\mathrm{var}(\theta | \mathcal{D}) = \frac{(a+N_{\mathrm{stab}})(b+N_{\mathrm{unstab}})}{(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}})^2(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}}+1)}
$$
Since for the chosen prior $a=b=1 \ll N$ this simplifies to
$$
\mathrm{var}(\theta | \mathcal{D}) = \frac{\hat{\theta}(1-\hat{\theta})}{N}
$$
yielding the error estimate given by the associated standard deviation. In all simulations we use $N~=~10000$ so that the maximum error for $\hat{\theta}~=~0.5$ is $\sigma~=~\sqrt{\mathrm{var}(\theta | \mathcal{D})} \approx 0.005$.
